A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-06-25 sections/Appendix.tex
Script statistics update; appendix table style
minor format: capitalizations, etc
camera-ready: acmart ver., fig. Description
pretty tables

        \centering\framebox{$|\Gm|_\to$, $|\jg|_\to$, $|A|_\to$} {Function measure.}
        \begin{gather*}
            \begin{aligned}
            |\Gm|_\to &= \sum_{\jg \in \Gm} |\jg|_\forall\\[2mm]
            |A \le B|_\forall &= |A|_\forall + |B|_\forall\\
            |e \Lto A|_\forall &= |A|_\forall\\
            |e \To \jg|_\forall &= |\jg|_\forall\\
            |\appInfAlg{A}{e}|_\forall &= |A|_\forall + |\jg|_\forall\\[2mm]
            |1|_\to &= |a|_\to = |\al|_\to = 1\\
            |A\to B|_\to &= |A|_\to + |B|_\to + 1\\
            |\all A|_\to &= |A|_\to
            % \sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
            %     \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\\
            \end{aligned}
        \end{gather*}
    \end{minipage}
\caption{Worklist measures}
\label{fig:measures}
\end{figure}

\subsection{Broken Lemmas in DK's papers}
\label{appendix:false_lemmas}

@@

>>>>

        \centering\framebox{$|\Gm|_\to$, $|\jg|_\to$, $|A|_\to$} {Function measure.}
        \begin{gather*}
            \begin{aligned}
            |\Gm|_\to &= \sum_{\jg \in \Gm} |\jg|_\forall\\[2mm]
            |A \le B|_\forall &= |A|_\forall + |B|_\forall\\
            |e \Lto A|_\forall &= |A|_\forall\\
            |e \To \jg|_\forall &= |\jg|_\forall\\
            |\appInfAlg{A}{e}|_\forall &= |A|_\forall + |\jg|_\forall\\[2mm]
            |1|_\to &= |a|_\to = |\al|_\to = 1\\
            |A\to B|_\to &= |A|_\to + |B|_\to + 1\\
            |\all A|_\to &= |A|_\to
            % \sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
            %     \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\\
            \end{aligned}
        \end{gather*}
    \end{minipage}
\Description{Worklist measures}
\caption{Worklist measures}
\label{fig:measures}
\end{figure}

\subsection{Broken Lemmas in DK's Papers}
\label{appendix:false_lemmas}

@@

\subsection{Translation Table for the Proof Scripts}
In the proof scripts, we use textual relational definitions
rather than the symbolic ones used in the paper.
The mapping should be helpful for anyone who reads the script.

\afterpage{%
\clearpage% Flush earlier floats (otherwise order might not be correct)
\begin{landscape}% Landscape page

\begin{table}[ht]
\renewcommand{\arraystretch}{1.3} 
\caption{Translation Table for the Proof Scripts}
\centering
\begin{tabular}{@{}llll@{}}
\toprule
Symbol & Abella textual relation & File (.thm) & Description\\
\midrule
$A$    & {\tt ty, wft, wfta} & typing &
    The "type" type, decl./alg. well-formedness\\
$e$    & {\tt tm, wftm} & typing &
    The "term" type, (alg.) well-formedness\\
$\Psi, \Gm$ & {\tt olist, wfj} & typing &
    {\tt olist} is Abella's built-in list type, i.e. {\tt [o]}\\
$\jg$  & {\tt judgment, wfjg} & typing &
    Judgment chain and its well-formedness\\

$\Gm\rto\Gm'$ & {\tt reduction} & smallStep &
    Algorithmic reduction: paper version\\
$\Gm\redto\nil$ & {\tt judge} & typing &
    ({\tt judge} $\Gm$): a success reduction on $\Gm$\\

$\Gm \sto \Om$ & {\tt tex} & trans &
    "transfer existential variables"\\
$\Om\redto\nil$ & {\tt dc, dcl} & trans, dcl &
    Declarative chain representation\\

$\Psi'\le\Psi$ & {\tt esub} & declTyping & Declarative context subtyping\\

$\Gm\jExt\Gm'$ & {\tt jExt} & inst\_decidable & "Judgment extension"\\
$\Gm_i$ & {\tt iexp} & inst\_decidable & instantiation judgments ($\Gm \vdash_{wf} \Gm_i$)\\
$\Gm_\le$ & {\tt subExp} & inst\_decidable & subtyping judgments\\

$|\cdot|_e$ & {\tt tmSize, tmSizeJ, tmSizel} & declTyping, decidability &
    Size measure for term, judgment chain or context\\
$|\cdot|_\Leftrightarrow$ & {\tt m\_judgeJ, m\_judge} & decidability &
    Judgment measure for judgment chain or context\\
$|\cdot|_{\al}$ & {\tt nVar} & inst\_decidable &
    Existential measure for context\\
$|\cdot|_\forall$ & {\tt order, orderJ, orderl} & order, inst\_decidable &
    Polymorphism measure for type, judgment chain or context\\
$|\cdot|_\to$ & {\tt depth, depthJ, depthl} & order, decidability &
    Function measure for type, judgment chain or context\\
\bottomrule
\end{tabular}
\label{table:translation_table}
\end{table}

\end{landscape}
\clearpage% Flush page
}

### 2019-06-25 sections/Metatheory.tex
camera-ready: acmart ver., fig. Description
pretty tables

\begin{figure}[t]
\begin{gather*}
\begin{aligned}
\text{Declarative worklist}\qquad&\Om &::=&\quad \nil \mid \Om, a \mid \Om, x: A \mid \Om \Vdash \jg
\end{aligned}
\end{gather*}
\hfill \framebox{$\Gm \sto \Om$} \hfill $\Gm$ instantiates to $\Om$.
\begin{gather*}
\inferrule*[right=$\mathtt{{\sto}}\Om$]
{~}
{\Om \sto \Om}
\quad
\inferrule*[right=$\mathtt{{\sto}\al}$]
{\Om\vdash\tau \\ \Om,[\tau/\al]\Gm \sto \Om}
{\Om,\al,\Gm \sto \Om}
\end{gather*}
\caption{Declarative Worklists and Instantiation}
\label{fig:trans}
\end{figure}

@@

\hfill \framebox{$\Om \rto \Om'$} \hfill Declarative transfer.
\begin{gather*}
\begin{aligned}
\Om,a &\rto \Om \\  \Om,x:A & \rto \Om\\
\Om\Vdash A\le B &\rto \Om &\text{ when } \|\Om\| \vdash A\le B\\
\Om\Vdash e\Lto A &\rto \Om & \text{ when } \|\Om\| \vdash e\Lto A\\
\Om\Vdash e\To_a \jg &\rto \Om\Vdash[A/a]\jg & \text{ when } \|\Om\| \vdash e\To A\\
\Om\Vdash \appInfAlg{A}{e} &\rto \Om\Vdash[C/a]\jg & \text{ when } \|\Om\| \vdash \appInf{A}{e}{C}\\
\end{aligned}
\end{gather*}
\caption{Declarative Transfer.}
\label{fig:decl:worklist}
\end{figure}

@@

\begin{figure}[t]
\framebox{$\Psi' \le \Psi$}
$$\begin{aligned}
\inferrule*[right=$\mathtt{CtxSubEmpty}$]
{~}{\nil \le \nil}\qquad
\inferrule*[right=$\mathtt{CtxSubTyVar}$]
{\Psi' \le \Psi}{\Psi', a \le \Psi, a}\qquad
\inferrule*[right=$\mathtt{CtxSubTmVar}$]
{\Psi' \le \Psi \\ \Psi \vdash A' \le A}{\Psi', x:A' \le \Psi, x:A}\qquad
\end{aligned}$$
\caption{Context Subtyping}
\label{fig:context_subtyping}
\end{figure}

@@

\begin{figure}
\hfill \framebox{$\Gm\jExt\Gm'$} \hfill \llap{$\Gm$ updates to $\Gm'$}.
\begin{gather*}
\inferrule*[right=$\mathtt{\jExt id}$]
    {~}{\Gm\jExt\Gm}
\qquad
\inferrule*[right=$\mathtt{\jExt solve}$]
    {|A|_\forall = 0 \quad \Gm_L,[A/\al]\Gm_R \jExt \Gm'}{\Gm_L,\al,\Gm_R \jExt \Gm'}
\qquad
\inferrule*[right=$\mathtt{\jExt \al}$]
    {\Gm_L,\al,\Gm_R \jExt \Gm'}{\Gm_L,\Gm_R \jExt \Gm'}
\end{gather*}
\caption{Worklist Update}\label{fig:worklist_ext}
\end{figure}

@@

\paragraph{Statistics of the Proof}
The proof script consists of 7977 lines of Abella code with a total of
60 definitions and 596 theorems.
Figure~\ref{table:proof_statistics} briefly summarizes the contents of each file.
The files are linearly dependent due to limitations of Abella.

\begin{table}[t]
\caption{Statistics for the proof scripts}
\centering\begin{tabular}{|c|c|c|l|}\hline
    File(s) & SLOC & \#Theorems & Description\\\hline
    olist.thm, nat.thm  &  311 & 57  & Basic data structures\\\hline
    typing.thm          &  245 & 7   & Declarative \& algorithmic system, debug examples\\\hline
    decl.thm            &  226 & 33  & Basic declarative properties\\\hline
    order.thm           &  235 & 27  & The $|\cdot|_\forall$ measure; decl. subtyping strengthening\\\hline
    alg.thm             &  679 & 80  & Basic algorithmic properties\\\hline
    trans.thm           &  616 & 53  & \makecell[l]{Worklist instantiation and declarative transfer\\
                            Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\\hline
    declTyping.thm      &  909 & 70  & \makecell[l]{Non-overlapping declarative system \\
                            Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll},
                                \ref{lem:inv_chkLam}, \ref{lem:subsumption}}\\\hline
    soundness.thm       & 1107 & 78  & Soundness theorem; aux. lemmas on transfer\\\hline
    depth.thm           &  206 & 14  & The $|\cdot|_\to$ measure; Lemma~\ref{lem:prune_inst}\\\hline
    dcl.thm             &  380 & 12  & Non-overlapping declarative worklist \\\hline
    completeness.thm    & 1124 & 61  &
                            Completeness theorem; aux. lemmas and relations\\\hline
    inst\_decidable.thm &  837 & 45  & Other worklist measures; Lemma~\ref{lemma:inst:decidable}\\\hline
    decidability.thm    &  983 & 57  & Decidability theorem and corollary\\\hline
    smallStep.thm       &  119 &  2  & The equivalence between big-step and small-step\\\hline
    \hline\emph{Total}               & 7977 & 596 & (60 definitions in total)\\\hline
\end{tabular}
\label{table:proof_statistics}
\end{table}

>>>>

\begin{figure}[t]
\begin{gather*}
\begin{aligned}
\text{Declarative worklist}\qquad&\Om &::=&\quad \nil \mid \Om, a \mid \Om, x: A \mid \Om \Vdash \jg
\end{aligned}
\end{gather*}
\hfill \framebox{$\Gm \sto \Om$} \hfill $\Gm$ instantiates to $\Om$.
\begin{gather*}
\inferrule*[right=$\mathtt{{\sto}}\Om$]
{~}
{\Om \sto \Om}
\quad
\inferrule*[right=$\mathtt{{\sto}\al}$]
{\Om\vdash\tau \\ \Om,[\tau/\al]\Gm \sto \Om}
{\Om,\al,\Gm \sto \Om}
\end{gather*}
\Description{Declarative Worklists and Instantiation}
\caption{Declarative Worklists and Instantiation}
\label{fig:trans}
\end{figure}

@@

\hfill \framebox{$\Om \rto \Om'$} \hfill Declarative transfer.
\begin{gather*}
\begin{aligned}
\Om,a &\rto \Om \\  \Om,x:A & \rto \Om\\
\Om\Vdash A\le B &\rto \Om &\text{ when } \|\Om\| \vdash A\le B\\
\Om\Vdash e\Lto A &\rto \Om & \text{ when } \|\Om\| \vdash e\Lto A\\
\Om\Vdash e\To_a \jg &\rto \Om\Vdash[A/a]\jg & \text{ when } \|\Om\| \vdash e\To A\\
\Om\Vdash \appInfAlg{A}{e} &\rto \Om\Vdash[C/a]\jg & \text{ when } \|\Om\| \vdash \appInf{A}{e}{C}\\
\end{aligned}
\end{gather*}
\Description{Declarative Transfer}
\caption{Declarative Transfer}
\label{fig:decl:worklist}
\end{figure}

@@

\begin{figure}[t]
\framebox{$\Psi' \le \Psi$}
$$\begin{aligned}
\inferrule*[right=$\mathtt{CtxSubEmpty}$]
{~}{\nil \le \nil}\qquad
\inferrule*[right=$\mathtt{CtxSubTyVar}$]
{\Psi' \le \Psi}{\Psi', a \le \Psi, a}\qquad
\inferrule*[right=$\mathtt{CtxSubTmVar}$]
{\Psi' \le \Psi \\ \Psi \vdash A' \le A}{\Psi', x:A' \le \Psi, x:A}\qquad
\end{aligned}$$
\Description{Context Subtyping}
\caption{Context Subtyping}
\label{fig:context_subtyping}
\end{figure}

@@

\begin{figure}
\hfill \framebox{$\Gm\jExt\Gm'$} \hfill \llap{$\Gm$ updates to $\Gm'$}.
\begin{gather*}
\inferrule*[right=$\mathtt{\jExt id}$]
    {~}{\Gm\jExt\Gm}
\qquad
\inferrule*[right=$\mathtt{\jExt solve}$]
    {|A|_\forall = 0 \quad \Gm_L,[A/\al]\Gm_R \jExt \Gm'}{\Gm_L,\al,\Gm_R \jExt \Gm'}
\qquad
\inferrule*[right=$\mathtt{\jExt \al}$]
    {\Gm_L,\al,\Gm_R \jExt \Gm'}{\Gm_L,\Gm_R \jExt \Gm'}
\end{gather*}
\Description{Worklist Update}
\caption{Worklist Update}\label{fig:worklist_ext}
\end{figure}

@@

\paragraph{Statistics of the Proof}
The proof script consists of 7,977 lines of Abella code with a total of
60 definitions and 596 theorems.
Figure~\ref{table:proof_statistics} briefly summarizes the contents of each file.
The files are linearly dependent due to limitations of Abella.

\begin{table}[t]
\renewcommand{\arraystretch}{1.2}
\caption{Statistics for the proof scripts}
\centering\begin{tabular}{@{}lrrl@{}}
\toprule
    File(s) & SLOC & \#Theorems & Description\\
\midrule
    olist.thm, nat.thm  &   311 & 57  & Basic data structures\\
    typing.thm          &   245 & 7   & Declarative \& algorithmic system, debug examples\\
    decl.thm            &   226 & 33  & Basic declarative properties\\
    order.thm           &   235 & 27  & The $|\cdot|_\forall$ measure; decl. subtyping strengthening\\
    alg.thm             &   679 & 80  & Basic algorithmic properties\\
    trans.thm           &   616 & 53  & \makecell[l]{Worklist instantiation and declarative transfer;\\
                            Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\
    declTyping.thm      &   909 & 70  & \makecell[l]{Non-overlapping declarative system; \\
                            Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll},
                                \ref{lem:inv_chkLam}, \ref{lem:subsumption}}\\
    soundness.thm       & 1,107 & 78  & Soundness theorem; aux. lemmas on transfer\\
    depth.thm           &   206 & 14  & The $|\cdot|_\to$ measure; Lemma~\ref{lem:prune_inst}\\
    dcl.thm             &   380 & 12  & Non-overlapping declarative worklist \\
    completeness.thm    & 1,124 & 61  &
                            Completeness theorem; aux. lemmas and relations\\
    inst\_decidable.thm &   837 & 45  & Other worklist measures; Lemma~\ref{lemma:inst:decidable}\\
    decidability.thm    &   983 & 57  & Decidability theorem and corollary\\
    smallStep.thm       &   119 &  2  & The equivalence between big-step and small-step\\
\midrule
    \emph{Total}        & 7,977 & 596 & (60 definitions in total)\\
\bottomrule
\end{tabular}
\label{table:proof_statistics}
\end{table}
